---
title: "a3_task4_Rammy_Salem"
author: "Rammy Salem"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# load packages
library(tidyverse)
library(janitor)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(lubridate)
```

```{r}
sea_urchins <- read_csv(here("data","urchins.csv")) # read urchins data
```

```{r}
purple_urchins_naples_jan2010 <- sea_urchins %>% # Start from sea urchins data
  mutate(DATE = mdy(DATE)) %>% # Convert the date to class “Date”
  mutate(YEAR = year(DATE)) %>% # Create a new column that only contains the year
  mutate(MONTH = month(DATE)) %>% # Create a new column that contains the month
  mutate(site_name = case_when( # Create a new column called site_full that contains the full name for each of the five sites
    SITE == "NAPL" ~ "Naples Reef",
    SITE == "AQUE" ~ "Arroyo Quemado Reef",
    SITE == "CARP" ~ "Carpinteria Reef",
    SITE == "MOHK" ~ "Mohawk Reef",
    SITE == "IVEE" ~ "Isla Vista Reef"
  )) %>%
  filter(COMMON_NAME == "Purple Urchin", 
         YEAR == 2010, 
         MONTH == 1, 
         site_name == "Naples Reef") # Filtered to only include observations for purple urchins
```

```{r}
# Look at and consider histograms and QQ-plots for the size distribution for each of the treatments (Control and Annual) to assess normality. 
#Histogram
ggplot(data = purple_urchins_naples_jan2010, aes(x = SIZE)) +
  geom_histogram(bins = 40) +
  facet_wrap(~TREATMENT) +
  labs(x = "Sea urchin diameter (cm), measured to the nearest 0.5 cm",
       y = "Number of Sea Urchins")
```

```{r}
ggplot(data = purple_urchins_naples_jan2010, aes(sample = SIZE)) +
         geom_qq() + 
         facet_wrap(~TREATMENT)
```

### The histograms and Q-Q plots suggest that both samples have a slight positive skew. 

```{r}
# Create a basic table of descriptive statistics for both groups (including at least the mean, standard deviation, and sample size)

purple_urchins_naples_jan2010_table <- purple_urchins_naples_jan2010 %>% # Start from purple urchins data
  group_by(TREATMENT) %>% # group by treatment
  summarize(mean_size = mean(SIZE),# Find mean size
            median_size = median(SIZE), # Find median size
            sd_size = sd(SIZE), # Find standard deviation size
            n = n()) %>% # Find sample size
  rename("Mean Size" = mean_size,
         "Median Size" = median_size,
         "SD" = sd_size) %>% 
  kbl() %>% # Use kableExtra to make table look nicer
  kable_styling() 
  
#View it
purple_urchins_naples_jan2010_table
```


```{r}
# Create vectors containing observations for each sample (TREATMENT)
annual_sample <- purple_urchins_naples_jan2010 %>% 
  filter(TREATMENT == "ANNUAL") %>% # filter for only annual treatment
  pull(SIZE) # create vector of size data

control_sample <- purple_urchins_naples_jan2010 %>%
  filter(TREATMENT == "CONTROL") %>% # filter only control treatment
  pull(SIZE) # Create vector of size data

t.test(x = annual_sample, control_sample) # Perform a two-sample t-test to compare means between the two groups
```
```

### The results of a two-sample t-test comparing the mean diameters of purple urchins reveal that, on average, the sizes of purple urchins in areas with annual kelp removal (*M* = 3.92, *SD* = .97) are not significantly greater than the sizes of purple urchins in areas in which giant kelp is left intact (*M* = 3.72, *SD* = .87),  *t*(117.66) = 1.15, *p* = .25. On average, purple urchins in areas with annual kelp removal were larger by approximately 0.20 cm (5.4%) than purple urchins in areas with no annual kelp removal, but this difference was not statistically significant. 

### The p-value of approximately .25 means that there is approximately a 25% chance that the two samples came from the same population, and a 25% possibility that the observed differences were due to chance alone as opposed to differences due to the treatment conditions, which is not enough evidence to reject the null hypothesis that the two samples are from the same population based on the the generally accepted below 5% *p*-value significance threshold.
